---
title: "Homework Assignment: bae-j-hwk-1-2"
output: pdf_document
date: "2023-01-25"
---

```{r setup, include=FALSE}

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, stringr, readxl, data.table, gdata)

full.ma.data <- readRDS("/Users/jenniferbae/Downloads/ECON 470 R Coding/homework1/data/output/full_ma_data.rds")
contract.services.area <- readRDS("/Users/jenniferbae/Downloads/ECON 470 R Coding/homework1/data/output/contract_service_area.rds")
ma.penetration.data <- readRDS("/Users/jenniferbae/Downloads/ECON 470 R Coding/homework1/data/output/ma_penetration.rds")
plan.premiums <- readRDS("/Users/jenniferbae/Downloads/ECON 470 R Coding/homework1/data/output/plan_premiums.rds")

```

```{r, include=FALSE}
load("Hwk1_workspace.Rdata")
```

# Enrollment Data

## 1. How many observations are in your current dataset?
There are 19126783 observations in my dataset full.ma.data


## 2. How many different plan_types are there?

```{r, include=FALSE}
# took out 1 from the total plan_types because one was categorized as "NA"

unique_type <- unique(full.ma.data$plan_type)
unique_type
```
There are 26 different plan_types in full.ma.data


## 3. Provide a table of the count of plans under each plan type in each year
```{r, echo=FALSE}

library(knitr)

plan.type.year1 <- full.ma.data %>%
  group_by(plan_type, year) %>%
  count() %>%
  arrange (year, -n) %>%
  filter (plan_type!= "NA")
plan.type.year1 <- pivot_wider(plan.type.year1, names_from = "year", values_from = "n")

knitr::kable (plan.type.year1)


```



## 4. Remove all special needs plans (SNP), employer group plans (eghp), and all “800-series” plans. Provide an updated version of Table 1 after making these exclusions.
```{r, echo=FALSE}

final.plans <- full.ma.data %>%
  filter(snp == "No" & eghp == "No" & (planid <800 | planid >= 900))

plan.type.year2 <- final.plans %>% 
  group_by(plan_type, year) %>%
  count () %>%
  arrange (year, -n)
plan.type.year2 <- pivot_wider(plan.type.year2, names_from = "year", values_from = "n")

knitr::kable (plan.type.year2)
```